Coupling finite and boundary element methods to solve the Poisson-Boltzmann equation for electrostatics in molecular solvation

### 2023-04-21 main.tex
Checked result section

@@

\title{Coupling finite and boundary element methods to solve the Poisson-Boltzmann equation for electrostatics in molecular solvation}
\author{Timo Betcke\thanks{Department of Mathematics, University College London, 25 Gordon Street, WC1H 0AY London, UK}, Micha\l{} Bosy\thanks{School of Computer Science and Mathematics, Kingston University London, Penrhyn Road, Kingston upon Thames, KT1 2EE, UK}, Erik Burman\thanks{Department of Mathematics, University College London, 25 Gordon Street, WC1H 0AY London, UK}, Christopher D. Cooper\thanks{Department of Mechanical Engineering and Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso, Universidad T\'ecnica Federico Santa Mar\'ia, Valpara\'iso 2390123, Chile}, Matthew W. Scroggs\thanks{Department of Mathematics, University College London, 25 Gordon Street, WC1H 0AY London, UK}}

>>>>

\usepackage{hyperref}

@@

\title{Coupling finite and boundary element methods to solve the Poisson-Boltzmann equation for electrostatics in molecular solvation}
\author{Timo Betcke\thanks{Department of Mathematics, University College London, 25 Gordon Street, WC1H 0AY London, UK}, Micha\l{} Bosy\thanks{School of Computer Science and Mathematics, Kingston University London, Penrhyn Road, Kingston upon Thames, KT1 2EE, UK}, Erik Burman\thanks{Department of Mathematics, University College London, 25 Gordon Street, WC1H 0AY London, UK}, \\ Christopher D. Cooper\thanks{Department of Mechanical Engineering and Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso, Universidad T\'ecnica Federico Santa Mar\'ia, Valpara\'iso 2390123, Chile}, Matthew W. Scroggs\thanks{Department of Mathematics, University College London, 25 Gordon Street, WC1H 0AY London, UK}}

### 2023-04-21 results.tex
Checked result section

For the finite element computations, we use the software package FEniCSx while for the boundary element computations we use Bempp-cl together with Exafmm-t. FEniCSx is the successor of the widely used FEniCS finite element library.
It provides a convenient Python interface, in which problems are described using UFL (Unified Form Language), a convenient domain-specific language specifically designed for finite element discretisations of partial differential equations. During assembly the UFL description is transformed into efficient low-level C++ code and just-in-time compiled. Bempp-cl is a Python package that uses low-level OpenCL kernels written in C99 to provide optimised assembly routines. The built-in dense assembly routines are highly efficient for moderate discretisation sizes up to a few ten thousand elements.

@@

To couple FEniCSx with Bempp-cl we load a volume mesh with FEniCSx. We then export the corresponding boundary mesh into Bempp-cl and assemble the boundary spaces there. Bempp-cl provides numerical trace operators that can translate from the degree of freedom (dof) representation in FEniCSx to the dof representation in Bempp-cl. The corresponding translation work is handled opaquely and the user only needs to deal with high-level interfaces of FEniCSx operators, Bempp-cl operators, and trace operators. FMM assembly fits automatically into this framework and can be enabled or disabled as simple configuration option.

Docker images containing FEniCSx, Bempp-cl, and Exafmm-t are publicly available (PROVIDE LINK), and all codes used in this sections are simple Jupyter Notebooks that can be reproducibly executed in this provided Docker image.

@@

The solvation energy computed with the two schemes is presented by Figure \ref{fig:arg_constant_energy}, which, as expected, converges to a similar answer as the mesh is refined.
Figure \ref{fig:arg2_constant_time_iter} compares the iteration count and time-to-solution. The left plot shows that using only BEM outperforms the coupled FEM-BEM approach in terms of iterations count. However, if we look at the total time that solvers take to obtain the solution, we can see the advantage of using the FEM-BEM coupling. The higher computational cost is caused by the need of using a hypersingular operator in the BEM formulation, and the fact that we are not using any acceleration method ($ie.$ FMM \textcolor{red}{true?}). Regardless, timings for FEM-BEM coupling scale much worse with size than its pure BEM counterpart, indicating that BEM-BEM coupling is better for larger problems. 
%\textcolor{red}{Chris: I thought BEM-BEM outperformed FEM-BEM in total time. Am I wrong there?} \textcolor{blue}{Michal: Not for this formulation of BEM-BEM since we have hypersingular operator. But you can see that for denser meshes we should outperform it.}

>>>>

For the finite element computations, we use the software package FEniCSx while for the boundary element computations, we use Bempp-cl together with Exafmm-t. FEniCSx is the successor of the widely used FEniCS finite element library.
It provides a convenient Python interface, in which problems are described using UFL (Unified Form Language), a convenient domain-specific language specifically designed for finite element discretisations of partial differential equations. During assembly, the UFL description is transformed into efficient low-level C++ code and just-in-time compiled. Bempp-cl is a Python package that uses low-level OpenCL kernels written in C99 to provide optimised assembly routines. The built-in dense assembly routines are highly efficient for moderate discretisation sizes up to a few ten thousand elements.

@@

To couple FEniCSx with Bempp-cl we load a volume mesh with FEniCSx. We then export the corresponding boundary mesh into Bempp-cl and assemble the boundary spaces there. Bempp-cl provides numerical trace operators that can translate from the degree of freedom (dof) representation in FEniCSx to the dof representation in Bempp-cl. The corresponding translation work is handled opaquely and the user only needs to deal with high-level interfaces of FEniCSx operators, Bempp-cl operators, and trace operators. FMM assembly fits automatically into this framework and can be enabled or disabled as a simple configuration option.

Docker images containing FEniCSx, Bempp-cl, and Exafmm-t are publicly available (\href{https://bempp.com/installation.html}{https://bempp.com/installation.html}), and all codes used in this sections are simple Jupyter Notebooks that can be reproducibly executed in this provided Docker image.

@@

The solvation energy computed with the two schemes is presented by Figure \ref{fig:arg_constant_energy}, which, as expected, converges to a similar answer as the mesh is refined.
Figure \ref{fig:arg2_constant_time_iter} compares the iteration count and time-to-solution. The left plot shows that using only BEM outperforms the coupled FEM-BEM approach in terms of iterations count. However, if we look at the total time that solvers take to obtain the solution, we can see the advantage of using the FEM-BEM coupling. The higher computational cost is caused by the need of using a hypersingular operator in the BEM formulation, and the fact that we are not using any acceleration method ($ie.$ FMM). Regardless, timings for FEM-BEM coupling scale much worse with size than its pure BEM counterpart, indicating that BEM-BEM coupling is better for larger problems. 
%\textcolor{red}{Chris: I thought BEM-BEM outperformed FEM-BEM in total time. Am I wrong there?} \textcolor{blue}{Michal: Not for this formulation of BEM-BEM since we have hypersingular operator. But you can see that for denser meshes we should outperform it.}
